---
title: "Lab 6"
author: "Daniel Varghese"
output: pdf_document
---

#Visualization with the package ggplot2


I highly recommend using the [ggplot cheat sheet](https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) as a reference resource. You will see questions that say "Create the best-looking plot". Among other things you may choose to do, remember to label the axes using real English, provide a title and subtitle. You may want to pick a theme and color scheme that you like and keep that constant throughout this lab. The default is fine if you are running short of time.

Load up the `GSSvocab` dataset in package `carData` as `X` and drop all observations with missing measurements. This will be a very hard visualization exercise since there is not a good model for vocab.

```{r}
[x] - 3

```

Briefly summarize the documentation on this dataset. What is the data type of each variable? What do you think is the response variable the collectors of this data had in mind?

The data type of each variable is a continous variable where you can have possible values that are very large or small. However, the nativeBorn category seems to be binary considering it is a yes or no. The response variables they had in mind are numbers or a yes or no answer for the nativeBorn category.

Create two different plots and identify the best-looking plot you can to examine the `age` variable. Save the best looking plot as an appropriately-named PDF.

```{r}
library(ggplot2)
ggplot2:(GGSvocab, x=age, y= nativeBorn)
```

Create two different plots and identify the best looking plot you can to examine the `vocab` variable. Save the best looking plot as an appropriately-named PDF.

```{r}
ggplot2:(GGSvocab, x=age, y= nativeBorn)
```

Create the best-looking plot you can to examine the `ageGroup` variable by `gender`. Does there appear to be an association? There are many ways to do this.

```{r}
library(ggplot2)
ggplot2(X) + 
    geom_jitter(aes(x = ageGroup, y = gender))
```

Create the best-looking plot you can to examine the `vocab` variable by `age`. Does there appear to be an association?

```{r}
ggplot(X) +
    geom_point(aes(x = vocab, y = age))
```

Add an estimate of $f(x)$ using the smoothing geometry to the previous plot. Does there appear to be an association now?

```{r}
ggplot(X) + 
  geom_jitter(aes(x = vocab, y = age)) +
  geom_smooth()
```

Using the plot from the previous question, create the best looking plot overloading with variable `gender`. Does there appear to be an interaction of `gender` and `age`?

```{r}
Yes I would say so.
```


Using the plot from the previous question, create the best looking plot overloading with variable `nativeBorn`. Does there appear to be an interaction of `nativeBorn` and `age`?

There seems to be an interaction between two people
#TO-DO
```

Create two different plots and identify the best-looking plot you can to examine the `vocab` variable by `educGroup`. Does there appear to be an association?

```{r}
There could be a correlation but not causation. 
```

Using the best-looking plot from the previous question, create the best looking overloading with variable `gender`. Does there appear to be an interaction of `gender` and `educGroup`?

```{r}
#There does seem to be an interaction.
```

Using facets, examine the relationship between `vocab` and `ageGroup`. You can drop year level `(Other)`. Are we getting dumber?

``{}
some people may be.
```
